---
title: "2020 Imnaha River Chinook Salmon and Bull Trout summaries"
author: "Joseph Feldhaus (ODFW) & Ryan Kinzer (NPT)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: console
---

## Noteable Highlights/Data issues  

Quick review of important events that occured during the 2020 trapping season.

  * Imnaha Weir installation date: **Friday, June 12, 2020 at 13:30 hrs**. 
  * PIT tag detections for Bull Trout and Chinook Salmon were limited to detection events that occured on or after 3/1/2020.
  * A tribal fishery occured below the weir and ended on *MONTH XX*, 2020.
  * The measuring board PIT tag antenna was not functioning on 6/14 and 6/15.  Therefore, 6 newly PIT tagged BT and 5 previously tagged (i.e., recaps) were not detected on IML antenna 9 which is node IMNAHWB0.
  * 6/25/2020: Chute #1 (next to panel that doesn’t open) was partially plugged with lodged sticks and rock.  Chutes #2 and #3 were open but did contain rocks in the chutes, but unable to determine if rocks were impeding passage.  All three chutes were cleared of rock and debris as of about 10 am on 6/25/2020.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning=FALSE)
```

```{r Source scripts, include=FALSE}
source("./R/02_load_ChinookDART_PTAGISbull_2020.R")
source("./R/03_process_Imnaha_data_2020.R")
```


```{r load_data, echo=FALSE, message=FALSE, warning=FALSE}
# Un-comment out the source code to update the data
#.libPaths("C:/R-3.6.1/library")
library(tidyverse)
library(lubridate)
library(pander)

yr <- 2020
 df<-readRDS("./data/PITcleanr_2020_chs_bull.rds")%>%
   mutate(Origin=t_rear_type)%>%
   filter(ObsDate>=ymd_hm(paste0(yr,"/03/01 00:00")))%>%droplevels()

detect_hist<-readRDS("./data/2020_detect_hist.rds")
```

## PIT summaries below the weir site
### Unique PIT-tag observations by species
```{r}
df %>%
  group_by(Mark.Species, Origin) %>%
  summarise(Unique_Tags = n_distinct(TagID)) %>%
  pander()
```

### Unique PIT-tag observations by species, origin and release site
```{r }
df %>%
  mutate(Release.Year = year(rel_date)) %>%
  group_by(Mark.Species, Origin, rel_site) %>%
  summarise(Unique_Tags = n_distinct(TagID)) %>%
  pander()
```



### Unique PIT-tag observations by species, SiteID and origin
Unexpanded PIT tag counts at Big Sheep Creek (BSC), and the mainstem PIT tag interrogation sites (IR1-IR5), the Imnaha adult ladder (IML), and the Imnaha facility trap house (IMNAHW).   
```{r }
df %>%
  mutate(Release.Year = year(rel_date)) %>%
  group_by(Mark.Species,SiteID,Origin) %>%
  summarise(Unique_Tags = n_distinct(TagID)) %>%
  pander()
```


### Unique PIT-tag observations within the Imnaha Basin by species, origin and arrival date
First detection dates of unexpanded Bull Trout and Chinook salmon PIT tags at tributary (COC, BSC) and mainstem (IR1-IR5, IML, IMNAHW) PIT tag interrogation sites.  Site are organized from downstream to upstream with the most downstream interrogation site (COC) in the top two panels.   
```{r Unique Tag counts, fig.width=8,fig.height=10,message=FALSE, warning=FALSE}

df %>%
  filter(!SiteID%in%c("BSC","COC")) %>%
  group_by(TagID, Mark.Species, Origin, rel_site, SiteID) %>%
   slice(which.min(lastObsDate)) %>%
  ggplot(aes(x=lastObsDate, fill=Origin)) +
  geom_histogram(colour = 'black', alpha = .5, bins = 100) +
  facet_grid(SiteID~Mark.Species,scales="free") +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw() +
  theme(legend.position = 'bottom') +
  theme(panel.grid.major=element_blank())+
  labs(x = "First detection date",
       y = "Count")
```

## PIT tag detection efficiency calculations
### Efficiency at the node level  

Detection efficiencies at the node level (i.e., upstream and downstream antenna grous) and conversion rates for Bull Trout and Chinook Salmon detected at IR1-IR5, IML, and IMNAHW.  Detections in the trap house from the measuring board PIT reader will be labeled "IMNAHWB0".

```{r efficiency}
library(PITcleanr)
load("./data/config_data_20190531.rda")

tmp_df <- df %>% 
  anti_join(detect_hist %>%
            filter(NewTag == 'TRUE') %>%
            select(TagID)
            ) %>%
  identity()

#tmp_df%>%filter(TagID=="3DD.00778C5775")


#names(tmp_df)

nodes <- c('IR1', 'IR2', 'IR3B0', 'IR3A0', 'IR4B0', 'IR4A0',
           'IMLB0', 'IMLA0', 'IMNAHWB0','IMNAHWA0','IR5B0', 'IR5A0')

tmp_df %>%
  group_by(Mark.Species) %>%
  do(estNodeEff(.,node_order = node_order, node = nodes)) %>%
    mutate(detLower95 = detEff - (1.96*detEff_SE),
         detUpper95 = detEff + (1.96*detEff_SE),
         tagsLower95 = estTagsAtNode - (1.96*estTags_SE),
         tagsUpper95 = estTagsAtNode + (1.96*estTags_SE)) %>%
  select(Mark.Species, Node, tagsAtNode, estTagsAtNode, tagsLower95, tagsUpper95,
         detEff, detLower95, detUpper95) %>%
  mutate(N_up = lag(estTagsAtNode),
         Conversion = paste0(round(estTagsAtNode/N_up*100),"%")) %>%
  select(-N_up) %>%
  pander()


```

### Efficiency at the site level  

Detection efficiencies and converson rates at the site level.  See the Node level estimates above for an efficiency estimate at IR5B0.
```{r echo=FALSE}

tmp_df <- df %>% 
  anti_join(detect_hist %>%
            filter(NewTag == 'TRUE') %>%
            select(TagID)
            ) %>%
  mutate(Node = gsub("A0", "B0", Node)) %>%# calculates efficiency at the site level
  identity()

nodes <- c('IR1', 'IR2', 'IR3B0', 'IR4B0',
           'IMLB0', 'IMNAHWB0', 'IR5B0')

tmp_df %>%
  group_by(Mark.Species) %>%
  do(estNodeEff(.,node_order = node_order, node = nodes)) %>%
  mutate(SiteID = gsub("B0", "", Node),
         detLower95 = detEff - (1.96*detEff_SE),
         detUpper95 = detEff + (1.96*detEff_SE),
         tagsLower95 = estTagsAtNode - (1.96*estTags_SE),
         tagsUpper95 = estTagsAtNode + (1.96*estTags_SE)) %>%
  select(Mark.Species, SiteID, tagsAtSite = tagsAtNode,
         estTagsAtSite = estTagsAtNode, tagsLower95, tagsUpper95,
         detEff, detLower95, detUpper95) %>%
 mutate(N_up = lag(estTagsAtSite),
         Conversion = paste0(round(estTagsAtSite/N_up*100),"%")) %>%
  select(-N_up) %>%
  pander()
```

## Summaries focused on the Weir   
### New tags vs Recaps at IMNAHW

```{r }

detect_hist %>%
  filter(Mark.Species=="Bull Trout",
         !is.na(IMNAHW))%>%
  select(Mark.Species,NewTag) %>%
  mutate(NewTag=ifelse(NewTag=="FALSE","New","Recap"))%>%
  group_by(Mark.Species, NewTag) %>%
  tally()%>%
  pivot_wider(names_from=NewTag,values_from=n)%>%
  mutate(Total=New+Recap)%>%
  pander()


```
### Unique PIT-tag observations at the Imnaha Weir by trap status (based on observation date at weir sites (IR4, IML, IMNAHW, IR5)

```{r }
detect_hist %>%
  select(Mark.Species, TrapStatus, IR4, IML, IMNAHW, IR5) %>%
  gather(SiteID, firstObsDate, IR4:IR5, na.rm = TRUE) %>%
  group_by(Mark.Species, TrapStatus, SiteID) %>%
  summarise(Unique_tags = n()) %>%
pander()  
```

### Passage routes of PIT-tagged fish successfully reaching and being detected at IR5 
This summary is limited to previously tagged fish (i.e., not tagged at IMNAHW in 2019) detected at IR5.  

*Passage Route descriptions:*  
* *Handled*: Processed in the trap house (IMNAHW) followed by a detection at IR5.  
* *IML obs = F *: Not detected at IML (F = False).  
* *IML obs = T*:  Detected at IML (T = True). 

*Trap Status descriptions*  
* *No obs at IR4*: no PIT tag observation at IR4.  
* *Panels Closed*: Indicates that the weir was fully operational.  
* *Panels Open*:  Separates out detections at IR5 before the weir was operational.  

 
```{r message=FALSE, warning=FALSE}
detect_hist %>%
  filter(grepl('Passed', TagStatus), NewTag == 'FALSE') %>%
  #filter(grepl('Passed', TagStatus)) %>%
  #group_by(Mark.Species, TrapStatus, PassageRoute) %>% 
  group_by(Mark.Species, PassageRoute, TrapStatus) %>%
  arrange(Mark.Species,PassageRoute)%>%
  summarise(Unique_tags = n()) %>%
pander()
```


### Unique PIT-tag observations by tag status  
*Tag status descriptions:*  
* **At Weir**: Observed at IR4 but not IML, IMNAHW, or IR5.   
* **Attempted Ladder**: Detected at IML, but not detected at IMNAHW.   
* **Last obs: BSC**: Observed at Big Sheep Creek.  
* **Last obs: IR1/IR2/IR3**: Observed at IR1, IR2, or IR3, respectively.  
* **NewTag**: Bull Trout tagged at IMNAHW in 2019.  
* **Passed**: Detected at IR5.  
* **Trapped**: Handled at IMNAHW, but no subsequent detection at IR5.  
* **Trapped: Obs Below Weir**: Handled at IMNAHW & detected at IR4>IMNAHW  
 


```{r}
detect_hist %>%
  group_by(Mark.Species, Origin=t_rear_type, TagStatus) %>%
  summarise(Unique_Tags = n()) %>%
  pander()

##filter for detections at IR4, IR5, IML, IMNAHW
AtWeir<-detect_hist%>%
  filter(str_detect(TagPath,"IR4|IR5|IML|IMNAHW"))%>%
  mutate(LastObsDate=as.Date(LastObs,format="%m/%d/%Y"))

ggplot(subset(AtWeir,LastObsDate>"2020/05/31"),aes(x=as.Date(LastObs,"%m/%d/%Y",tz="UTC"),fill=TagStatus))+
  geom_histogram(colour = 'black', alpha = .5, binwidth=1)+
  geom_vline(xintercept = as.numeric(as.Date("2020-06-12")), linetype=2)+
  facet_wrap(~Mark.Species,scales="free_y",ncol=1)+
  scale_x_date(date_breaks="5 days",date_label="%m-%d")+
  scale_fill_brewer(palette = 'Set1') +
  theme_bw() +
  theme(legend.position = 'bottom')+ 
  guides(fill=guide_legend(nrow=2,byrow=TRUE))+
  theme(panel.grid.major=element_blank())+
  ggtitle("Tag status: last detection at IR4/IR5/IML/IMNAHW on or after weir installation")+
  labs(x = "Last Observation Date",
       y = "Count")
```


### Tag Status = Last observed at IR4 just below the weir

Fish assigned a tag status of "At Weir" have been detected at IR4 but have not been detected at IML, handled in the trap (IMNAHW), or detected at IR5. IR4_min and IR4_max are the first and last detection dates at IR4, respectively.  The *fl_last* represents that last recorded fork length (mm), either at tagging or as a recapture. For Chinook salmon, lengths less than 350 mm were assumed to be lengths from juvenile tagging events and were set to NA.   

```{r Tag Status At Weir, echo=FALSE, message=FALSE, warning=FALSE}
TagStatusAtWeir<-detect_hist %>%
  filter(TagStatus=="Last Obs: IR4") %>%
  select(TagID, 
         Species = Mark.Species,
         fl_last=fl_max,
         rel_site,
         Origin=t_rear_type,
         TagStatus,
         IR4,IR4_max) %>%
  arrange(Species,IR4_max)

TagStatusAtWeir %>%
  group_by(Species,Origin) %>%
  tally() %>%
  pander()

TagStatusAtWeir %>%
  rename(IR4_min =IR4,
         Release.Site = rel_site) %>%
  mutate(IR4_min = as.Date(IR4_min,format="%m/%d"),
         IR4_max=as.Date(IR4_max,format="%m/%d")) %>%
  select(-TagStatus,-Origin) %>%
  pander()

```

### Tag Status = "Passed"

List of PIT tags detected at IR5 were assigned a tag status of passed.  
```{r}
TagStatusPassed<-detect_hist %>%
  filter(TagStatus=="Passed") %>%
  select(TagID, 
         Species = Mark.Species,
         fl_last=fl_max,
         rel_site,
         Origin=t_rear_type,
         TagStatus,
         IR5,
         IR5_max) %>%
  arrange(Species,IR5_max)

TagStatusPassed %>%
  group_by(Species,Origin) %>%
  tally() %>%
  pander()

TagStatusPassed %>%
  rename(IR5_min =IR5,
         Release.Site = rel_site) %>%
  mutate(IR5_min = as.Date(IR5_min,format="%m/%d"),
         IR5_max=as.Date(IR5_max,format="%m/%d")) %>%
  select(-TagStatus,-Origin) %>%
  pander()
```



## Travel time between sites

```{r Travel Time, fig.width=8,fig.height=10}
detect_hist %>%
  select(TagID, Mark.Species, Origin=t_rear_type, Arrival_Month, IR1_IR3:IR4_IR5) %>%
  mutate_at(c(4:6), as.numeric) %>%
  gather(Reach, Travel_Time, IR1_IR3:IR4_IR5) %>%
  mutate(Reach = fct_relevel(Reach, c("IR1_IR3", "IR3_IR4", "IR4_IML", "IML_IMNAHW", "IR4_IMNAHW", "IR4_IR5"))) %>%
  filter(Travel_Time >= 0) %>%
  #filter(Reach == "IR4_IR5") %>%
  ggplot(aes(x = Travel_Time, fill = Origin)) +
  geom_histogram(colour = 'black', alpha = .5, binwidth = 1) +
  scale_fill_brewer(palette = 'Set1') +
  facet_grid(Reach ~ Mark.Species, scales = "free") +
  theme_bw() +
  theme(legend.position = 'bottom') +
  labs(x = 'Travel Time (Days)',
       y = 'Count') 
```



```{r days_site, fig.width=8,fig.height=10, eval = FALSE}
### Number of days spent at a site
# number of days at a site
df %>%
  filter(!SiteID %in% c('COC','BSC')) %>%
  group_by(Mark.Species, Origin, SiteID, TagID) %>%
  summarise(minObsDate = min(lastObsDate),
          maxObsDate = max(lastObsDate),
          hours = difftime(maxObsDate, minObsDate, units = 'days')) %>%
  ggplot(aes(x = hours, fill = Origin)) +
  geom_histogram(colour = 'black', alpha = .5, bins = 20) +
  scale_fill_brewer(palette = 'Set1') +
  facet_grid(SiteID ~ Mark.Species) +
  theme_bw() +
  theme(legend.position = 'bottom') +
  labs(x = 'Days (max(obsDate) - min(obsDate))',
       y = 'Count') 
```

### Tag Status = "Last Obs:  IR3" 

Summary of tags last observed at IR3. Some of these tags will "drop-out" before reaching the weir site (IR4/IML/IR5).

```{r Tag Status At IR3, echo=FALSE, message=FALSE, warning=FALSE}
IR3tags<-detect_hist%>%
  filter(TagStatus=="Last obs: IR3")%>%
  select(TagID,Mark.Species,t_rear_type,TagStatus,IR3)%>%
  arrange(Mark.Species,IR3)

IR3tags%>%group_by(Mark.Species,t_rear_type)%>%
  tally() %>% pander()

#IR3tags %>% rename(IR3_min=IR3)%>%mutate(IR3_min=as.Date(IR3_min,format="%m/%d"))%>% select(-TagStatus)%>%pander()


```


```{r, eval = FALSE}
###Minimum Travel Time from IR4 to IMNAHW  

detect_hist %>%
  select(TagID, Mark.Species, Origin=t_rear_type, IR4_IML,IR4_IMNAHW,IML_IMNAHW) %>%
  mutate_at(c(4:6), as.numeric) %>%
  gather(Reach, Travel_Time, IR4_IML,IR4_IMNAHW,IML_IMNAHW) %>%
  filter(Travel_Time >= 0) %>%
  mutate(Reach=factor(Reach,levels=c('IR4_IML','IR4_IMNAHW','IML_IMNAHW')))%>%
  #filter(Reach == "IR4_IR5") %>%
  ggplot(aes(x = Travel_Time, fill = Origin)) +
  geom_histogram(colour = 'black', alpha = .5, bins = 20) +
  scale_fill_brewer(palette = 'Set1') +
  facet_grid(Reach ~ Mark.Species, scales = "free") +
  theme_bw() +
  theme(legend.position = 'bottom') +
  labs(x = 'Travel Time (Days)',
       y = 'Count') 
  
```




### Source Data
The complete PIT-tag histories for Chinook Salmon and Bull trout detected during 2020 in the Imnaha River Basin were downloaded as two seperate files.  The Chinook detections were downloaded from the Columbia River DART website [DART](http://www.cbr.washington.edu/dart/cs/data/nezperce/nptspawn_GRA_2020_1.csv). The Bull Trout detections are the result of running a "Complete Tag History" parameterized on the [PTAGIS website](https://www.ptagis.org/) and set-up to run and save automatically at 7:00 a.m. and 2:00 p.m. each day.  The Bull trout detections are stored within the "*feldhauj/2020BullTroutCompleteTagHistory.csv*" filepath.  We are then using an R-script to download and combine the files from the DART website and the PTAGIS ftp server into a single file.  Once the file is combined all Chinook Salmon and Bull trout tag detections are processed using the R package [PITcleanR](https://github.com/kevinsee/PITcleanR). The final dataset provides a simplified tag history with the first and last detection dates at each PIT-tag detection node (i.e., a single spanning array or antenna) and information regarding upstream and downstream movements and a more general migration direction. 


### *Creating the PTAGIS Bull Trout Complete Tag History report:*  
* *Query:* All PIT-tagged Bull trout detected in the Imnaha River Basin.   
* *Event Date:* Between 3/1/2020 and 12/31/2020. For this report, limited to dates after 31 March.
* *Event Sites:* IR1, IR2, IR3, IR4, IR5, IML, IMNAHW, BSC, IMNAHR
* *Mark Species:* Bull Trout
* *Event Type:* Observation, Recapture, Recovery 


### *Output Files:*  
Some very brief meta data descriptions for two output files processed with R scripts.  For more details on these files, contact Joseph Feldhaus (Joseph.Feldhaus@state.or.us) or Ryan Kinzer (ryank@nezperce.org).

  * *PITcleanr_2020_chs_bull.xlsx*:  The complete tag histories from the Chinook DART detections and the PTAGIS Bull Trout detections combined into a single file and processed with the PITcleanr R-package.  

  * *detect_hist.xlsx*: A pivot table style summary. (This description is not complete JF 6/22/2020) Each row = a unique PIT tag code.  Columns correspond to first detection dates at IR1-IR5, IML, and IMNAHW.  This file contains the Trap Status, TagPath, Passage Route, and TagStatus fields.  Trap status references the dates the weir was operating. TagPath is a character string representing detections at PIT tag observations sites.  TagStatus represents the last known location of the tag and whether the tag has arrived at the weir, has attempted the ladder (i.e., detected at IML), passed the weir (i.e., detected at IR5), or has been trapped (i.e., detected at IMNAHW). The tag pathway describes the passage route through the weir.
